iSEE: Interactive SummarizedExperiment Explorer

### 2018-05-05 manuscript/main.tex
Correct affiliations CS

\title{iSEE: Interactive SummarizedExperiment Explorer}
\author[1,$\dagger$]{Kevin Rue-Albrecht}
\author[2,3,$\dagger$]{Federico Marini}
\author[4,$\dagger$]{Charlotte Soneson}
\author[5,$\dagger$,$\ast$]{Aaron T. L. Lun}

\affil[1]{Kennedy Institute of Rheumatology, University of Oxford, Oxford OX3 7FY, United Kingdom}
\affil[2]{Center for Thrombosis and Hemostasis (CTH), University Medical Center of the Johannes Gutenberg University Mainz, Mainz}
\affil[3]{Institute for Medical Biostatistics, Epidemiology and Informatics (IMBEI), University Medical Center of the Johannes Gutenberg University Mainz, Mainz}
\affil[4]{University of Zurich and SIB Swiss Institute of Bioinformatics, Zurich, Switzerland}
\affil[5]{Cancer Research UK Cambridge Institute, University of Cambridge, Cambridge CB2 0RE, United Kingdom}
\affil[$\dagger$]{These authors contributed equally.}
\affil[$\ast$]{To whom correspondence should be addressed.}

@@

\begin{itemize}
\item Column data plots, for visualizing sample metadata stored in the colData slot of SummarizedExperiment object.
\item Feature expression plots, for visualizing expression values for a particular feature (e.g., gene) across samples from any assay in the SummarizedExperiment object.
Despite its name, this plot type is applicable to data modalities beyond gene expression.
\item Row statistics tables, to present the contents of the rowData slot for the SummarizedExperiment object.
\item Row data plots, for visualizing sample metadata stored in the rowData slot of a SummarizedExperiment object.
\item Heat maps, to visualize assay data for multiple features where samples are ordered by one or more colData fields.
\item Reduced dimension plots, which display any two dimensions from pre-computed dimensionality reduction results (e.g., from PCA or $t$-SNE).
These results are taken from the reducedDim slot of any SingleCellExperiment object supplied to iSEE.
\end{itemize}

\begin{figure*}[t]
\includegraphics[width=\textwidth]{pics/Fig1.pdf}
\caption{iSEE uses a customisable multi-panel layout (A) that simultaneously displays one or more panels of various types, where each panel type visualizes a different aspect of the data.
New panels of any type can be added (i), and all panels can be removed, reordered or resized (ii).
Panel types are available to visualize sample-based reduced dimensionality embeddings (iii), sample-level metadata (iv), and experimental observations across samples for each feature (v).
Other panel types include row statistics tables (vi), to facilitate searching across features and their metadata; heatmaps (vii), to visualize experimental observations for multiple features; and feature-level metadata plots.
Panels of each type are colour-coded for ease of interpretation.
(B) Information can be transmitted between panels according to a user-specified scheme.
Here, the selection of feature $X$ in the row statistics table determines the y-axis of the feature expression plot, and colours the samples in the reduced dimension plot by the expression of $X$.
Selection of points in the reduced dimension plot (dotted blue line) also determines the samples that are shown in the column data (i.e., sample metadata) plot;
further selection of points in the column data plot determines the samples that are shown in the heatmap.
}
\label{fig:iSEE}
\end{figure*}

Each sample is represented as a point in column data, feature expression and reduced dimension plots.
Similarly, each feature is represented by a point in row data plots.
For these panel types, a scatter plot is automatically produced if the selected variables on the x- and y-axes are both continuous.
If exactly one variable is categorical, points are grouped by the categorical levels and a (horizontal) violin plot is produced with points scattered within each violin.
If both variables are categorical, a ``rectangle plot'' is produced where each combination of categorical levels is represented by a rectangle with area proportional to the frequency of that combination.
Points are scattered randomly within each rectangle.
For ease of interpretation, the rectangle plot collapses to a mirrored bar plot when one of the categorical variables only has one level.

@@

\subsection*{Dynamic linking between panels}
A key feature of iSEE is the ability to dynamically transmit information between panels (Fig.~\ref{fig:iSEE}B).
Users can define and reorganize arbitrary links between ``transmitting'' and ``receiving'' panels, whereby selections in transmitting panels control the inclusion and appearance of the corresponding data points in receiving panels.
This feature facilitates exploration of the relationships between different aspects of the data.
For example, users can easily determine co-expression patterns of genes in a particular region of a reduced dimensionality embedding -- this is achieved by selecting points in a reduced dimension plot (using the standard rectangular brush or a lasso) and transmitting that selection to any number of feature expression plots.\\

>>>>

\title{iSEE: Interactive SummarizedExperiment Explorer}
\author[1,$\dagger$]{Kevin Rue-Albrecht}
\author[2,3,$\dagger$]{Federico Marini}
\author[4,5,$\dagger$]{Charlotte Soneson}
\author[6,$\dagger$,$\ast$]{Aaron T. L. Lun}

\affil[1]{Kennedy Institute of Rheumatology, University of Oxford, Oxford OX3 7FY, United Kingdom}
\affil[2]{Center for Thrombosis and Hemostasis (CTH), University Medical Center of the Johannes Gutenberg University Mainz, Mainz}
\affil[3]{Institute for Medical Biostatistics, Epidemiology and Informatics (IMBEI), University Medical Center of the Johannes Gutenberg University Mainz, Mainz}
\affil[4]{Institute of Molecular Life Sciences, University of Zurich, 8057 Zurich, Switzerland}
\affil[5]{SIB Swiss Institute of Bioinformatics, University of Zurich, 8057 Zurich, Switzerland}
\affil[6]{Cancer Research UK Cambridge Institute, University of Cambridge, Cambridge CB2 0RE, United Kingdom}
\affil[$\dagger$]{These authors contributed equally.}
\affil[$\ast$]{To whom correspondence should be addressed.}

@@

\begin{itemize}
\item Column data plots, for visualizing sample metadata stored in the colData slot of SummarizedExperiment object.
\item Feature assay plots, for visualizing expression values for a particular feature (e.g., gene) across samples from any assay in the SummarizedExperiment object.
Despite its name, this plot type is applicable to data modalities beyond gene expression.
\item Row statistics tables, to present the contents of the rowData slot for the SummarizedExperiment object.
\item Row data plots, for visualizing sample metadata stored in the rowData slot of a SummarizedExperiment object.
\item Heat maps, to visualize assay data for multiple features where samples are ordered by one or more colData fields.
\item Reduced dimension plots, which display any two dimensions from pre-computed dimensionality reduction results (e.g., from PCA or $t$-SNE).
These results are taken from the reducedDim slot of any SingleCellExperiment object supplied to iSEE.
\end{itemize}

\begin{figure*}[t]
\includegraphics[width=\textwidth]{pics/Fig1.pdf}
\caption{iSEE uses a customisable multi-panel layout (A) that simultaneously displays one or more panels of various types, where each panel type visualizes a different aspect of the data.
New panels of any type can be added (i), and all panels can be removed, reordered or resized (ii).
Panel types are available to visualize sample-based reduced dimensionality embeddings (iii), sample-level metadata (iv), and experimental observations across samples for each feature (v).
Other panel types include row statistics tables (vi), to facilitate searching across features and their metadata; heatmaps (vii), to visualize experimental observations for multiple features; and feature-level metadata plots.
Panels of each type are colour-coded for ease of interpretation.
(B) Information can be transmitted between panels according to a user-specified scheme.
Here, the selection of feature $X$ in the row statistics table determines the y-axis of the feature assay plot, and colours the samples in the reduced dimension plot by the expression of $X$.
Selection of points in the reduced dimension plot (dotted blue line) also determines the samples that are shown in the column data (i.e., sample metadata) plot;
further selection of points in the column data plot determines the samples that are shown in the heatmap.
}
\label{fig:iSEE}
\end{figure*}

Each sample is represented as a point in column data, feature assay and reduced dimension plots.
Similarly, each feature is represented by a point in row data plots.
For these panel types, a scatter plot is automatically produced if the selected variables on the x- and y-axes are both continuous.
If exactly one variable is categorical, points are grouped by the categorical levels and a (horizontal) violin plot is produced with points scattered within each violin.
If both variables are categorical, a ``rectangle plot'' is produced where each combination of categorical levels is represented by a rectangle with area proportional to the frequency of that combination.
Points are scattered randomly within each rectangle.
For ease of interpretation, the rectangle plot collapses to a mirrored bar plot when one of the categorical variables only has one level.

@@

\subsection*{Dynamic linking between panels}
A key feature of iSEE is the ability to dynamically transmit information between panels (Fig.~\ref{fig:iSEE}B).
Users can define and reorganize arbitrary links between ``transmitting'' and ``receiving'' panels, whereby selections in transmitting panels control the inclusion and appearance of the corresponding data points in receiving panels.
This feature facilitates exploration of the relationships between different aspects of the data.
For example, users can easily determine co-expression patterns of genes in a particular region of a reduced dimensionality embedding -- this is achieved by selecting points in a reduced dimension plot (using the standard rectangular brush or a lasso) and transmitting that selection to any number of feature assay plots.\\
